---
title: "Week_1(Diet_data_case_study)"
author: "Priyadarshini De"
date: "January 30, 2019"
output: html_document
---
# Setting directory information
```{r}
path <- "C:/Users/depri/Desktop/advanced data analytics/R_and_descriptive_analytics_review(week1)/Diet Case Study"
setwd(path)
getwd()
```
# printing my working directory
```{r}
my_wd <- getwd()
print(my_wd)
```
# Listing the directory and or files
# list.dirs these functions produce a character vector of the names of files or directories in the named directory.
```{r}
list.dirs() # List the data directories
files <- list.files("diet_data")  # list the data files
files

```
# Basic understanding and basic statistics of the data, # look into one data file
# read.csv reads a file in the table format and creates a dataframe from it with cases corresponding to lines and #variable to fields in the file.
```{r}
andy = read.csv("diet_data/Andy.csv")
head(andy) # examine the first 6 rows in the file
length(andy$Day) # this is a vector
dim(andy)  # read the dimension of the data file
str(andy)  # Compactly display an internal structure of the R object
```
# Read column headers
```{r}
names(andy)
```
# Get basic statistics about the data
```{r}
summary(andy)
```
# Subsetting the data
# View the weight of andy on day1
# Subset the data , the first row of the weight column
```{r}
andy[1, "Weight"] 
# View the weight of andy on day 10th
andy[10, "Weight"]
```
# Understanding the various approaches to filer data
# filter rows based on a dynamic condition method 1 use the square brackets as follows #dataframe[row_filter,column_filter]
```{r}
andy[andy$Day==30, "Weight"] # Or we can use the subset function
subset(andy$Weight,andy$Day==30) # or we can use the which function that returns the row index of the specified condition/filter
andy[which(andy$Day==30), "Weight"]

```
# Extracting useful information 
# How much weight andy lost from day 1 till Day 30?
```{r}
andy_start_date = andy[1,"Weight"]
andy_end_date = andy[30, "Weight"]
andy_lost_weight = andy_start_date - andy_end_date
print(andy_lost_weight)
```
# Reading data from another file (John data file)
```{r}
#head(read.csv(files[3]))
```
# what is wrong incorrect path missing folder name how can we fix this issue?
# Reading the filenames as well as their containing folder
```{r}
data_files_full = list.files("diet_data", full.names = TRUE)
data_files_full
```
```{r}
head(read.csv(data_files_full[3]))
```
# Binding the data frame of andy and david(binding two data frames)
```{r}
andy_david = rbind(andy, read.csv(data_files_full[2]))
head(andy_david)
tail(andy_david)
```
# Selecting certain rows from the binded data
```{r}
which(andy_david$Day==25)
```
# Selecting certain rows from the binded data
```{r}
day_25 = andy_david[which(andy_david$Day==25),] # or,
day_25 = andy_david[andy_david$Day==30,]
print(day_25) # or we can solve the reuslt for future use
criteria = which(andy_david$Day==25)
andy_david[criteria,]
```
# or selecting rows from the binded data
```{r}
andy_david[which(andy_david$Day==25),]
```
# Append everybody's data to the same dataframe
# Manually using loops and rbind
```{r}
dat = data.frame() # create an empty dataframe called "dat" to hold the binded data
for(i in 1:5){
  dat = rbind(dat,read.csv(data_files_full[i]))
}
str(dat)
```
# Calculating the median of the weight column
```{r}
median(dat$Weight)
```
# It contains na lets see which row has na
```{r}
print(dat[77,])
```
# Getting the median of the weight column
```{r}
median(dat$Weight, na.rm=TRUE)
```
# Calculating the median of day 30 in all the data collection
```{r}
subjects_day_30 = dat[which(dat[,"Day"]==30),]
print(subjects_day_30)

```
# median
```{r}
median(subjects_day_30$Weight)
```
# We have done a lot of manual data manipulation so far let us build a function that will return the median weight #of a given day
```{r}
weight_median = function(directory, day) {
list_files= list.files(directory, full.names = TRUE) # create a list of files
dat = data.frame()     # create an empty data frame
for (i in 1:5)          # Loop through the file rbinding them together
dat = rbind(dat, read.csv(list_files[i]))



dat_subset = dat[which(dat[, "Day"]==day),] # Subsets the rows that match the day argument

median(dat_subset[, "Weight"], na.rm=TRUE) # gettin the median of the column weight

}
```
# Testing the function using various ways
```{r}
weight_median(directory="diet_data", day=20) # or
weight_median("diet_data", 20)
weight_median("diet_data", 4)
weight_median("diet_data", 17)
```
# The better approach is to create an output object of an appropriate size and then fill it up
```{r}
files_full = list.files("diet_data", full.names=TRUE)
summary(files_full)
```
# Create an empty list that's the length of our expected output.
# Our input object is "files_full" and our empty list is "going to be tmp"
```{r}
tmp = vector(mode = "list", length = length(files_full))
summary(tmp)
```
# Now we need to read in those csv files and drop them into our tmp list. Read in each of the csv files and place them inside our list
```{r}
for (i in seq_along(files_full)){
  tmp[[i]] = read.csv(files_full[[i]])  # each element in the tmp list contains a dataframe of a patient file.
}
str(tmp)
```
# what we did was read in each of the csv files and place them inside our list.
# Now we have a list of 5 elements called tmp where each element of the list is a data frame containing one of the csv files. This is functionally identical to using lapply
```{r}
tmp1 = lapply(files_full, read.csv)
str(tmp1)
```
# or, 
```{r}
str(lapply(files_full, read.csv))
```
# Now how can we look at the data
```{r}
str(tmp[[1]])
```
# Lets look at the first 6 observations
```{r}
head(tmp[[1]][,"Day"])
```
# To manually rbind everything together
```{r}
output1 = rbind(tmp[[1]], tmp[[2]], tmp[[3]],tmp[[4]], tmp[[5]])
str(output1)
```
# Better way use do.call()
# The syntax is do.call(function you want to use, list of arguments)
# do.call executes a function on a list of arguments to be passed to it.In our case we want to rbind() our list of data frames which tmp in this case
```{r}

output = do.call(rbind, tmp)
str(output)
```



























